Fail Faster

### 2025-03-12 paper/paper.tex
scala overhead

\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  Bq.bind (Bq.gen_int 0 100) (fun x ->
    Bq.bind (Bq.gen_int 0 x) (fun y ->
      Bq.return (x,y)
  ))

@@

\jwc{Demonstrate this in Scala too!!!}

@@

\jwc{
  \begin{itemize}
    \item Because the RNG sits at the heart of the generation hot path, RNG
    speed matters enormously. Which RNG you pick matters...

    \item Different PBT libraries use different RNGs --- QuickCheck and \bq use
    splittable_random, Scalacheck uses \href{https://burtleburtle.net/bob/rand/smallprng.html}{this one}.  These
    are reasonably performant RNGs, but faster ones exist. Because this is the
    core of the hot path of the generator, speeding up sampling just a little can
    can dramatically speed up generation time.
    \item Significantly faster RNGs that still pass pRNG benchmarks like BigCrush exist (like \href{https://github.com/lemire/testingRNG/blob/master/source/lehmer64.h}{this one}), and using them can 
    speed up generation enormously.

    \item However, showing this
    is tricky, because different RNGs have different sampling orders, but our
    eval is made much stronger by the fact that all of our generators are functionally equivalent.

    \item In this paper demonstrate this fact by noticing a ``natural experiment'': OCaml's splittable random library is
    slow in a way that can be improved \emph{without} changing the extensional behavior. By resolving this issue, we can demonstrate
    just how much you can improve bug-finding performance simply by speeding up the RNG.

    % To illustrate this, we demonstrate a particular inefficiency of
    % OCaml's BQ, resolve it in a later section, and then show how much you can
    % bug-finding performance just by speeding up and RNG.
  \end{itemize}
}

>>>>

\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  let%bind x = (Bq.gen_int 0 100) in
  let%bind y = (Bq.gen_int 0 x) in
  Bq.return (x,y)

@@

\jwc{Same story in scala. Approximately 2x difference for the same generator... (scala uses a slightly different generator type, but the principles are the same)}

\begin{lstlisting}
def intPair : Gen[(Long,Long)] = for {
  x <- Gen.choose(0,1000)
  y <- Gen.choose(0,x)
} yield (x,y)

def intPairInlined : (Gen.Parameters, Seed) => (Option[(Long,Long)],Seed) = {
 (p,seed) =>
  val (x,seed2) = chLng(0,1000)(p,seed)
  x match {
    case None => (None,seed2)
    case Some(x) =>
      val (y,seed3) = chLng(0,x)(p,seed2)
      y match {
        case None => (None,seed)
        case Some(y) => (Some(x,y),seed3)
      }
    }
}
\end{lstlisting}

% [info] GenBm.generateComplexBase     avgt    3  458.325 ± 67.214  ns/op
% [info] GenBm.generateComplexInlined  avgt    3  266.035 ±  8.559  ns/op

@@

\jwc{
  \begin{itemize}
    \item Because the RNG sits at the heart of the generation hot path, RNG
    speed matters enormously. Which RNG you pick matters...
    \item Different PBT libraries use different RNGs --- QuickCheck and \bq use
    splittable random, Scalacheck uses \href{https://burtleburtle.net/bob/rand/smallprng.html}{this one}.  These
    are reasonably performant RNGs, but faster ones exist. Because this is the
    core of the hot path of the generator, speeding up sampling just a little can
    can dramatically speed up generation time.
    \item Significantly faster RNGs that still pass pRNG benchmarks like BigCrush exist (like \href{https://github.com/lemire/testingRNG/blob/master/source/lehmer64.h}{this one}), and using them can 
    speed up generation enormously.
    \item However, showing this is tricky, because different RNGs have different sampling orders, but our
    eval is made much stronger by the fact that all of our generators are functionally equivalent.
    \item In this paper demonstrate this fact by noticing a ``natural experiment'': OCaml's splittable random library is
    slow in a way that can be improved \emph{without} changing the extensional behavior. By resolving this issue, we can demonstrate
    just how much you can improve bug-finding performance simply by speeding up the RNG.
  \end{itemize}
}
